Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-03-22 WSSSPE2_report/WSSSPE2_review.tex
Section 3.2: Add some content.

>>>>

In a recent survey of UK research-intensive universities, 92\% of researchers
said they used research software and 68\% said their research would be impossible
without software. Yet 71\% have had no formal software training, and few are
ready to apply many of the things we take for granted such as testing or
virtualisation. WSSSPE represents the pinnacle of what we understand to be the
best practice around scientific software in our community. My talk will
challenge the workshop participants to come up with ways of taking this best
practice to those 92\% of researchers in a way that will lead to maximum benefit
to the scientific community.

%\url{http://dx.doi.org/10.6084/m9.ﬁgshare.1243288}

Where indicated slides licensed under World-leading research relies on software

``Today there are very few science areas left which do not rely on IT and thus
software for the majority of their research work. More importantly key
scientific advances in experimental and observational science would have been
impossible without better software.'' Kersten Kleese van Dam Pacific Northwestern
National Laboratory via change.org campaign

``Scientific discovery and innovation are advancing along fundamentally new
pathways opened by development of increasingly sophisticated software. Software
is an integral enabler of computation, experiment and theory, and directly
responsible for increased scientific productivity and enhancement of
researchers' capabilities.'' Dan Katz SI2 Program Director, National Science
Foundation.

We do not recognise these hundreds of thousands of researchers because we think
they are the long tail and forget that they are the mainstream and we are the
elite minority

Software is not special it is mainstream software?

What would happen to your research without software

Survey of researchers from 15 Russell Group unis conducted by SSI between Aug- Oct 2014.
406 respondents covering representative range of funders, discipline and seniority.

This is not just about the ``traditional'' sciences

And 	attitudes 	are	changing

Overall

 which 	is	changing 	the	situation

PhD students

Survey of  PhD students from Russell Group university conducted by SSI between Aug - Sep 2014.
~100 respondents covering representative range of discipline.

Even	though 	a	lot	is	still	the	same

Survey of 138 University professionals conducted by SSI in Feb-Apr 2013.

The cost of UK research that relies on software is 
£840m!

Investment in 2013-2014 financial year, an amount that has risen by
3\% on average over last four years

30\%

Of total research investment has been spent on research which relies on software
over the last four financial years

Analysis of data from 49,650 grant titles and abstracts published on Gateway to
Research covering 2010-2014.

The UK research community is just starting to understand the magnitude of the issue

56\%
Of UK researchers develop their
own research software

71\%
Of UK researchers have had no formal
software development training

140,000
UK researchers are relying on
their own coding skills

4\%
Of jobs advertised in UK
universities were software related

Survey of researchers from 15 Russell Group unis conducted by SSI between Aug - Oct 2014.

If 92\% of researchers care about software and WSSSPE is about caring about
software---what are we going to do to help researchers?

Sustainability

With apologies to Greg Wilson:
\url{http://software-carpentry.org/blog/2013/10/you-keep-using-that-word.html}

We	can 	talk	all	day 	about 	the best 	ways	of	achieving sustainability

But	what	are	we	going 	to	do to	make	it	actually sustainable?

Careers outside academic sector

Non-university Research (industry, government etc.)

UK	STEM graduate career paths

Early Career
Research

Permanent
Research Staff

Professor

\begin{verbatim}
Source: The Scientific Century, Royal Society, 2010 (revised to reﬂect ﬁrst stage
clariﬁcaFon from ``What Do PhD’s Do?'' study)
\end{verbatim}

Who	are	the	people 	we	need to	persuade?

What		are	the	incentives 	we need 	to	put 	in	place?

How are we going to show that the work we present here has benefit to the 92\%?

We	are	all	the 92\%	and	we can	change	the world	if	we	want!

\url{http://dx.doi.org/10.6084/m9.ﬁgshare.1243288}
